---
title: "KN-NBA"
author: "Bennett Apostol, Tate Davison, Emily Lantz, and Donny Robbins"
date: '2022-05-10'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
packages <- c("knitr", "ggplot2", "tidyverse", "tidymodels", 
              "ISLR", "caret", "class","glue","kknn")
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

## Presentation Overview

This is R Markdown document describes the our group final project outline for MATH/COSC 3570: Introduction to Data Science! RMarkdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. 

However, we plan to highlight the more complex functions of R, using concepts like **K-Nearest Neighbors (KNN)** and **Linear Regression**, and incorporating them into data from the NBA Playoffs. 

More importantly, for more details on the data-sets we will be using in our Final Project, click on the following links below:
- <https://www.basketball-reference.com/leagues/NBA_2022.html#all_shooting_teamopponent>
- (Insert Basketball csv's here)

## Data Frame Findings (Donny)
```{r prevData,echo = FALSE}
prev_year_data <- read_csv("prev_year_data.csv")
print(prev_year_data)
```
- Bullet 1
- Bullet 2
- Bullet 3
## Slide 2 (Emily Lantz)

```{r}
#summary(data_rk$Age)
#hist(data_rk$Age)
#data_rk_df <- as.data.frame(data_rk)
#data_rk_matrix<-as.matrix(data_rk)
#plot(frequency ~ Age, data = data_rk_matrix)



```

## Slide 3 (Insert Name)

```{r pressure}
plot(pressure)
```


## KNN Algorithm: Finding ideal K (Tate)

```{r, include=TRUE}
## Define training and test data

## Choose which variables to use

## Find ideal K as done in HW3


```


## KNN Algorithm Continued (Tate)

```{r, include=TRUE}

## Train Model using ideal K from previous slide

## Find test error rate, evaluate performance, perhaps make ROC curve 

## Make prediction using model 

```

## (MAYBE) Principle Component Analysis (Tate)

```{r, include=TRUE}
## Subset data to desired variables, perform PCA, obtain outputs

## Provide interpretations of PC1 and PC2, (I envision this being defensive and offensive aptitude)

## Make biplot of PC1 and PC2 and add vectors 

## Make prediction using model

```




