---
title: "KN-NBA"
author: "Bennett Apostol, Tate Davison, Emily Lantz, and Donny Robbins"
date: '2022-05-10'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
packages <- c("knitr", "ggplot2", "tidyverse", "tidymodels", 
              "ISLR", "caret", "class","glue","kknn")
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

tidy_data <- function(file){
  data <- read_csv(file)
  #Help from https://stackoverflow.com/questions/22104962/how-to-remove-empty-columns-in-r/27098520
  
  data <- Filter(function(x)!all(is.na(x)), data) #remove empty columns
  colnames(data)[18] <- "OeFG%"
  colnames(data)[19] <- "OTOV%"
  colnames(data)[21] <- "OFT/FGA"
  colnames(data)[22] <- "DeFG%"
  colnames(data)[23] <- "DTOV"
  colnames(data)[25] <- "DFT/FGA"
  data <- data[-c(1,26,27)]
  
  asc_method <- c("L","PL","OTOV%","DeFG%","DFT/FGA","DRtg")
  desc_method <- c("W","PW","MOV","SOS","SRS","ORtg","Pace","FTr",
                   "3PAr","TS%","OeFG%","ORB%","OFT/FGA","DTOV","Attend./G",
                   "NRtg","DRB%","Age")
  for (i in asc_method){
    data[i] <- rank(data[i],ties.method = "min")
  }
  for (i in desc_method){
    data[i] <- rank(-data[i],ties.method = "min")
  }
  return (data[-c(4,6)]) #removes Losses and PL
}
data_rk <- tidy_data("basketball data.csv")

data_rk <- data_rk %>% filter(endsWith(Team,"*")) %>% #filters out nonplayoff 
  select(everything()) %>%                            #teams
  mutate(Team = str_replace(Team,"\\*",""))

data_20 <- tidy_data("NBA20-21 data.csv")
data_19 <- tidy_data("NBA19-20 data.csv")
data_18 <- tidy_data("NBA18-19 data.csv")
data_17 <- tidy_data("NBA17-18 data.csv")
data_16 <- tidy_data("NBA16-17 data.csv")
data_15 <- tidy_data("NBA15-16 data.csv")
data_14 <- tidy_data("NBA14-15 data.csv")
data_13 <- tidy_data("NBA13-14 data.csv")
data_12 <- tidy_data("NBA12-13 data.csv")
data_11 <- tidy_data("NBA11-12 data.csv")
data_10 <- tidy_data("NBA10-11 data.csv")
data_09 <- tidy_data("NBA09-10 data.csv")
data_08 <- tidy_data("NBA08-09 data.csv")
data_07 <- tidy_data("NBA07-08 data.csv")
data_06 <- tidy_data("NBA06-07 data.csv")
data_05 <- tidy_data("NBA05-06 data.csv")
data_04 <- tidy_data("NBA04-05 data.csv")
data_03 <- tidy_data("NBA03-04 data.csv")
data_02 <- tidy_data("NBA02-03 data.csv")
data_01 <- tidy_data("NBA01-02 data.csv")
data_00 <- tidy_data("NBA00-01 data.csv")


prev_year_data <- data_20 %>% filter(endsWith(Team,"!")) %>%
  select(everything())
find_winner <- function(data){
  year_data <- data %>% filter(endsWith(Team,"!")) %>%
    select(everything())
  prev_year_data <- rbind(prev_year_data,year_data)
  return(prev_year_data)
}
prev_frames <- list(data_19,data_18,data_17,data_16,data_15,data_14,
                   data_13,data_12,data_11,data_10,data_09,
                   data_08,data_07,data_06,data_05,data_04,
                   data_03,data_02,data_01,data_00)

for (i in prev_frames){
  prev_year_data <- find_winner(i)
}

prev_year_data <- prev_year_data %>% mutate(Team = str_replace(Team,"\\!",""))
summarized <- summarize_all(prev_year_data,mean)


prev_year_data <- rbind(prev_year_data,summarized) %>%  
  mutate_if(is.numeric, ~round(., 0))
prev_year_data$Team[22] = "Average"


prev_year_data_csv<-write_csv(prev_year_data, "prev_year_data_csv")

```

## Presentation Overview (Bennett)
This is R Markdown document describes our group final project for MATH/COSC 3570: Introduction to Data Science — ***The NBA Playoffs!*** We examined the winners of the NBA Playoffs from the last 21 years, and observed if there were any trends or correlations between the winning teams.

We plan to highlight the more complex functions of R, using concepts like **K-Nearest Neighbors (KNN)** and **Linear Regression**, and incorporating them into data from the NBA Playoffs. 

For more details on the data-sets we will be using in our Final Project, click on the following links below:
<https://www.basketball-reference.com/leagues/NBA_2022.html#all_shooting_teamopponent>

## Wins vs. Simple Rating System
```{r, echo = FALSE}
SRS_plot <- ggplot(prev_year_data, aes(x = W, y = SRS)) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm") + 
  labs(x = "Wins",
       y = "Simple Rating System",
       title = "Wins vs. Simple Rating System")
SRS_plot
```

## Correlation Coefficient for W vs. SRS
```{r, echo = TRUE}
x = prev_year_data$W 
y = prev_year_data$SRS 
z <- cor(x, y)

print(z)
```

## Data Frame Findings (Donny)
```{r prevData,echo = FALSE}
prev_year_data <- read_csv("prev_year_data.csv")
print(prev_year_data)
```

## Slide 2 (Emily Lantz)
```{r,echo = TRUE}
summary(prev_year_data_csv$Age)
```
## Slide 3 (Emily Lantz)
```{r,echo = TRUE}
summary(prev_year_data$W)
```
## Slide 4 (Emily Lantz)
```{r,echo = TRUE}
hist(prev_year_data_csv$Age, breaks = 30)
```
## Slide 5 (Emily Lantz)
```{r,echo = TRUE}
plot(W ~ Age, data = prev_year_data_csv)
```
## Slide 6 (Emily Lantz)
```{r,echo = TRUE}
freq_Age_lm <- lm(W ~ Age, data=prev_year_data_csv)
```
## Slide 7 (Emily Lantz)
```{r,echo = FALSE}
#creating graph about the linear regression
ageGraph<-ggplot(prev_year_data_csv, aes(x=Age, y=W))+
  geom_point()+
  geom_smooth(method="lm", col="black")+
  theme_bw() +
  labs(title = "NBA Winners vs. Age of Team",
       x = "Age",
       y = "Wins")
ageGraph

```

## Slide 8 (Emily Lantz)
```{r,echo = TRUE}
ggplot(data=prev_year_data_csv, aes(x=W, y=Age))+ geom_point(alpha=0.5)+
  geom_smooth(method="lm", fill="yellow", se=TRUE)
```


## Slide 10 (Tate)

- *K-Nearest-Neighbors* 

- Predicting Probability of winning a championship using 22 variables

- 10 seasons worth of data on every NBA team

- Classification based on probability
```{r, echo=FALSE}

## Defining train and test data
knn_trn <- data_10 %>%
  rbind(data_11) %>%
  rbind(data_12) %>%
  rbind(data_13) %>%
  rbind(data_14) %>%
  rbind(data_16) %>%
  rbind(data_17) %>%
  rbind(data_18)

knn_tst <- data_15 %>%
  rbind(data_19)

## Adding logical column: true if the team won the title
knn_trn <- endsWith(x = knn_trn$Team, suffix = "!") %>%
  cbind(knn_trn)

knn_trn <- rename(knn_trn, "champ" = .)
knn_trn$champ <- as_factor(knn_trn$champ)


knn_tst <- endsWith(x = knn_tst$Team, suffix = "!") %>%
  cbind(knn_tst)

knn_tst <- rename(knn_tst, "champ" = .)
knn_tst$champ <- as_factor(knn_tst$champ)

```


## Slide 11 (Tate)

```{r, include=FALSE}

## Create and train model
knn_mdl <- parsnip::nearest_neighbor(neighbors = 3) %>%
  set_mode("classification") %>%
  set_engine("kknn")

knn_fit <- knn_mdl %>%
  fit(champ ~ ., data = select(knn_trn, -c("Team")))


## Performance on test data

knn_results <- bind_cols(
  knn_tst$Team,
  knn_tst$champ,
  predict(knn_fit, select(knn_tst, -c("Team"))),
  predict(knn_fit, select(knn_tst, -c("Team")), 
          type = "prob")
)

knn_results <- rename(knn_results, "Champ" = ...2, "Team" = ...1)


knn_pred <- pull(
  predict(knn_fit, select(knn_tst, -c("Team"))))

```
- Prediction results on test data: 
```{r, echo=FALSE}
table(knn_pred, knn_tst$champ)
```

- Test error rate
```{r, echo=FALSE}
mean(knn_pred != knn_tst$champ)
```

## Slide 12 (Tate)

- Correctly predicted Cavaliers to win the 2015-2016 finals 
```{r, echo=FALSE}
knn_results[2:5,]
```
- Incorrectly predicted Raptors would win the 2019-2020 finals
```{r, echo=FALSE}
knn_results[52,]

```

