---
title: "KN-NBA"
author: "Bennett Apostol, Tate Davidson, Emily Lantz, and Donny Robbins"
date: '2022-05-10'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Presentation Overview

This is R Markdown document describes the our group final project outline for MATH/COSC 3570: Introduction to Data Science! RMarkdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. 

However, we plan to highlight the more complex functions of R, using concepts like **K-Nearest Neighbors (KNN)** and **Linear Regression**, and incorporating them into data from the NBA Playoffs. 

More importantly, for more details on the data-sets we will be using in our Final Project, click on the following links below:
- <https://www.basketball-reference.com/leagues/NBA_2022.html#all_shooting_teamopponent>
- (Insert Basketball csv's here)

## Slide 1 (Insert Name)

- Bullet 1
- Bullet 2
- Bullet 3

## Slide 2 (Insert Name)

```{r cars, echo = TRUE}
summary(cars)
```

## Slide 3 (Insert Name)

```{r pressure}
plot(pressure)
```


## Slide 4 (Insert Name)

- Text Block 1
- Text Block 2
- Text Block 3

## Slide 5 (Insert Name)

- Insert Text
- Insert Text